Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-05-09 exp1/iswc2013.tex
32run in CI
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@544 43ef1190-90fd-4dd7-a769-d0121f02b80e
relax discussion of non-linearity
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@545 43ef1190-90fd-4dd7-a769-d0121f02b80e

The above experimental setup is executed across a design matrix
generated with the DoE.base package, which returns a data frame
consisting of the factors with a column containing the corresponding
curl measurements. To analyze the experiment, we first generate a
normal plot, see figure~\ref{fig:fullnormal}\todo{explain axes}. In this plot, if the
null hypothesis $H_0$ holds, all estimated effects will fall on a
straight line, as their means would be zero. Any significant departure
from a straight line can be interpreted as a significant effect (see
Section~4.8 of \cite{wu2009experiments} for a detailed
explanation). In our case, it is most important to note that any
effect that is negative (i.e. has a negative slope) means the effect
enhances the performance of the SPARQL endpoint, the runtime
decreases. If ``Implement'' and the sum of its interactions are
negative, then it supports the alternative hypothesis, i.e. we have
successfully improved the performance.

@@

We are forced to conclude that the three preceding experiments are
severely flawed, as they failed to take into account the non-linear
effects arising from larger data sizes on slow hardware.

@@

We saw that while 2 level experiments can perform well in determining
significant effects, they do not necessarily work well to find a
model, as exemplified by failure to identify the probable non-linear
effect of the P-ATA disks. This issue would be fixed by using an
orthogonal array design.

>>>>

The above experimental setup is executed across a design matrix
generated with the DoE.base package, which returns a data frame
consisting of the factors with a column containing the corresponding
curl measurements. To analyze the experiment, we first generate a
normal plot, see figure~\ref{fig:fullnormal}\todo{explain axes}. In this plot, if the
null hypothesis $H_0$ holds, all estimated effects will fall on a
straight line, as their means would be zero. Any significant departure
from a straight line can be interpreted as a significant effect (see
Section~4.8 of \cite{wu2009experiments} for a detailed
explanation). In our case, it is most important to note that any
effect that is negative means the effect
enhances the performance of the SPARQL endpoint, the runtime
decreases. If ``Implement'' and the sum of its interactions are
negative, then it supports the alternative hypothesis, i.e. we have
successfully improved the performance.

@@

We are forced to conclude that the three preceding experiments are
severely flawed, as they failed to take into account the effects from
larger data sizes on slow hardware.

@@

We saw that while 2 level experiments can perform well in determining
significant effects, they do not necessarily work well to find a
model, as exemplified by failure to identify the possibly non-linear
effect of the P-ATA disks. This issue would be fixed by using an
orthogonal array design.

Also, experiments as small as the 32-run is not useful in estimating
any detailed characteristics and is thus of little use for science,
but could be useful in engineering: If the experiment has been
validated by a larger experiment, it could serve well in a Continuous
Integration system.
